Learning from Data Through Models

### 2023-08-29 Paper/draft.Rmd
reshape simulation data to merge with data

---
title: "Learning from Data Through Models"
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Alberto Bisin, Guillaume Frechette and Jimena Galindo"
abstract: "TBW"
#classoption: pagebackref # passes pagebackref=true to hyperref before it is loaded, allowing backreferencing
output: 
    pdf_document:
        citation_package: natbib
        fig_width: 7
        fig_height: 6
        fig_caption: true
        number_sections: true

@@

bibliography: bibliography.bib
fontsize: 12pt
geometry: margin=2.5cm
---

```{r dependencies, include=FALSE}
source("../parameters_and_packages.R")
```

@@

# Literature Review

@@

After observing the aoutcome of each gamble, the agent updates their beliefs of both theta and omega in one of 4 differente ways: 
As a \emph{dogmatic} as in \cite ; as a \emph{switcher} as in \cite ; as a \emph{self-attribution} updater as in \cite ; 
or as a fully Bayesian agent. Each of these updating rules is described in detail in what follows.

>>>>

---
title: "Learning from Data Through Models"
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Jimena Galindo"
abstract: "TBW"
#classoption: pagebackref # passes pagebackref=true to hyperref before it is loaded, allowing backreferencing
output: 
    pdf_document:
        citation_package: natbib
        fig_width: 7
        fig_height: 6
        fig_caption: true
        number_sections: true

@@

bibliography: references.bib
fontsize: 12pt
geometry: margin=2.5cm
---

```{r dependencies, include=FALSE}
source("../Paper/parameters_and_packages.R")
```

@@

@@

After observing the outcome of each gamble, the agent updates their beliefs of both theta and omega in one of 4 differente ways: 
As a \emph{dogmatic} as in @Heidhues2018 ; as a \emph{switcher} as in @Ba2023 ; as a \emph{self-attribution} updater as in @Coutts2020; 
or as a fully Bayesian agent. Each of these updating rules is described in detail below.

Notice that both \theta and \omega can be identified from the outcomes.
However, under the assumption that the agent is myopic, the choices might not gererate sufficient variation in effort to 
identify both. Such cases arise when the agent is either overconfident or underconfident about their type and the combination
of their choices with the exogenous state are such that the data observed is consistent with the erroneous beliefs. Each of 
the updating mecahnisms above will make different predicitons about the cases in which the agent is able to learn their true 
type and when the misspecification persists.

## The Bayesian Benchmark

A fully rational agent updates their beliefs about $\theta$ and $\omega$ simultaneously by using Bayes' rule. 
The agent's posterior beliefs about $\theta$ and $\omega$ after observing a success are given by:
\begin{equation}
p[\theta_j|\text{success}] = \frac{p[\text{success}|\theta_j]p[\theta_j]}{\sum_{k=1}^3 p[\text{success}|\theta_k]p[\theta_k]}
\end{equation}
and
\begin{equation}
p[\omega_i|\text{success}] = \frac{p[\text{success}|\omega_i]p[\omega_i]}{\sum_{k=1}^3 p[\text{success}|\omega_k]p[\omega_k]}
\end{equation}

where $p[\text{success}|\theta_j]$ is the probability of success conditional on type $\theta_j$ and $p[\text{success}|\omega_i]$ is the
probability of success conditional on state $\omega_i$.

Notice that since agents are myopic, even though all the parameters could be identified with enough variation in choices, even 
a fully Bayesian agent might not learn their true type. This is because they do not internalize the tradeoff between flow payoff
and learning.
